---
title: "My answers"
author: "My name"
date: "2021-10-22"
output: html_document
---




> This little tutorial aims to make you familiar with some of the functions of the `stringr` package and a few regular expressions. 



## Strings and escape sequences  in `r fontawesome::fa(name = "r-project", fill = "steelblue")`

##### Write a sentence with escape sequences.
Try the sentence: `"It's the end of the world!" he said.\` . 
Assign the string to a variable and try as`print()`, `cat()` and `writeLines()`.

```{r}
# Write your answer here
```


## `stringr` functions 

We will be using the `words` data that is built into `stringr`. 
The data set is availble to you if you load the package.

```{r}
length(words)
```


##### Select words that

1. Contain a `y`
```{r}
# Write your answer here
```


2. Start with `y`
```{r}
# Write your answer here
```



3. Contain a `y` within the word

```{r}
# Write your answer here
```


##### Extract the `y` and the previous character.

Note: Use the function `unique()` around the results to avoid printing many empty matches.

```{r}
# Write your answer here
```



##### Get the lengths of the first ten words

You can use `head(words, 10)` as a convenient way to access the elements of the `words` vector.
```{r}
# Write your answer here
```



## Viral research 

Read the genome sequence of the Hepatitis D virus: [hepd.fasta](https://biostat2.uni.lu/practicals/data/hepd.fasta).
For now, just execute the following:

```{r}
hepd <- readr::read_lines("https://biostat2.uni.lu/practicals/data/hepd.fasta")
```

##### What is the length of the genome sequence? `r fontawesome::fa("ruler")`

```{r}
# Write your answer here
```


##### What is the sequence composition? How often does each character occur?

```{r}
# Write your answer here
```



##### Find motifs in the sequence using `str_locate()`. 

Find all matches of the sequence $ATG$ in the sequence. 

```{r}
# Write your answer here
```



