---
title: "My answers"
author: "My name"
date: "2022-05-11"
output: html_document
---




This guided practical will demonstrate that the **tidyverse** allows to compute summary statistics and visualize datasets efficiently.
This dataset is already stored in a **tidy** `tibble`, cleaning steps will come in future practicals.


##### Those kind of questions are optional {.bonus}

## `datasauRus` package

- check if you have the package `datasauRus` installed

```{r}
library(datasauRus)
```

- should return nothing. If `there is no package called ‘datasauRus’` appears, it means that the package needs 
to be installed. Use this:


## Explore the dataset

Since we are dealing with a `tibble`, we can type



only the first **10** rows are displayed.


##### What are the dimensions of this dataset? Rows and columns?

- **base** version, using either `dim()`, `ncol()` and `nrow()`

```{r}
# Write your answer here
```


- **tidyverse** version

```{r}
# Write your answer here
```


##### Assign the `datasaurus_dozen` to the `ds_dozen` name. This aims at populating the **Global Environment**

```{r}
# Write your answer here
```


##### Using Rstudio, those dimensions are now also reported within the interface, where?

```{r}
# Write your answer here
```


## How many datasets are present?

- **base** version

```{block, opts.label = "tip"}
You want to count the number of **unique** elements in the column **dataset**. The character `$` applied to a data.frame subset the column and convert the 2D structure to 1D, _i. e_ a vector.
The function **length()** returns the length of a vector, such as the unique elements
```

```{r}
# Write your answer here
```


- **tidyverse** version

```{r}
# n_distinct counts the unique elements in a given vector.
# we use summarise to return only the desired column named n here.
# we use English verbs and no subsetting characters, nor we change dimensions (keep a tibble)
summarise(ds_dozen, n = n_distinct(dataset))
```

- even better, compute and display the number of lines per `dataset`

```{block, opts.label = "tip"}
the function `count` in `dplyr` does the `group_by()` by the specified column + `summarise(n = n())` which returns the number of observation per defined group.
```

```{r}
# Write your answer here
```



## Check summary statistics per dataset

##### Compute the mean of the `x` & `y` column. For this, you need to `group_by()` the appropriate column and then `summarise()`

```{block, opts.label = "tip"}
in `summarise()` you can define as many new columns as you wish. No need to call it for every single variable.
```

```{r}
# Write your answer here
```


```{r}
# Write your answer here
```



##### Compute both mean and standard deviation (sd) in one go using `across()` {.bonus}


```{r}
# Write your answer here
```


```{r}
# Write your answer here
```



##### What can you conclude?

```{r}
# Write your answer here
```


## Plot the _datasauRus_

##### Plot the `ds_dozen` with `ggplot` such the **aesthetics** are `aes(x = x, y = y)` 
with the **geometry** `geom_point()`

```{block, opts.label = "tip"}
the `ggplot()` and `geom_point()` functions must be linked with a **+** sign
```

```{r}
# Write your answer here
```


##### Reuse the above command, and now colored by the `dataset` column

```{r}
# Write your answer here
```


Too many datasets are displayed.

##### How can we plot only one at a time?

```{block, opts.label = "tip"}
You can filter for one dataset upstream of plotting
```

```{r}
# Write your answer here
```


##### Adjust the **filtering** step to plot two datasets

```{block, opts.label = "tip"}
R provides the inline instruction `%in%` to test if there a match in the left operand with the right one (a **vector** most probably)
```

```{r}
# Write your answer here
```


##### Expand now by getting one `dataset` per **facet**

```{r}
# Write your answer here
```


##### Remove the filtering step to facet all datasets

```{r}
# Write your answer here
```



##### Tweak the theme and use the `theme_void` and remove the legend

```{r}
# Write your answer here
```


##### Are the datasets actually that similar?

```{r}
# Write your answer here
```



## Animation

Plots can be animated, see for example what can be done with [`gganimate`](https://gganimate.com/).
Instead of panels, **states** are made across `datasets` and **transitions** smoothed with an afterglow effect.

```{r}
# Write your answer here
```


## Conclusion

> Never trust summary statistics alone; always visualize your data | Alberto Cairo

**Authors**

- Alberto Cairo, (creator)
- Justin Matejka
- George Fitzmaurice
- Lucy McGowan

from this [post](https://itsalocke.com/datasaurus-now-cran/)
