---
title: "My answers"
author: "My name"
date: "2021-10-22"
output: html_document
---






> In this practical, you'll learn how to import flat files using the `readr` package


##### Those kind of questions are optional {.bonus}

# Before you start


To perform reproducible research it is a good practice to store the files in a standardized location. For example, you could take advantage of the RStudio projects and store data files in a sub-folder called `data`.

If you did not create an Rstudio project yet, **create a RStudio project** now.

##### Prepare your project's folder

1. Check that the project is active: the name you chose should appear on the top-right corner.
    
2. Create a folder named `data` within your project's folder. Use the `Files` pane in the lower right Rstudio panel or your favorite file browser.

3. Download the file [`blood_fat.csv`](https://biostat2.uni.lu/practicals/data/blood_fat.csv) and place it in the `data` sub-folder you just created.


4. Create a new **Rmarkdown** file, save it at the project root with a relevant name.

5. Add a code chunk and with those lines to load the libraries. You don't need to install the packages if those lines are working fine

~~~
library(dplyr)
library(readr)
~~~


6. Don't forget to run the chunk's code to load the library during your interactive session

```{block, opts.label = "warning"}
If you load the library only in the console and forget to place a chunk to load it, the knitting process will fail. Indeed, when you click on the `knit` button, the chunks are evaluated in a new and fresh environment.
```

# Use `readr` to load your first file


##### Load the `blood_fat` file

```{block, opts.label = "tip"}
the relative path can be safely built using `"data/blood_fat.csv"` 
if you followed the preliminary steps above, download the csv in a sub-folder `data` of a RStudio project

For example, you folder structure could be (depending on the picked names). Here:

- RStudio project is `Rworkshop`
- Rmarkdown document is `practical02_import.Rmd`

~~~
.
├── data
│   └── blood_fat.csv
├── practical03_import.Rmd
└── Rworkshop.Rproj
~~~

```


```{r}
# Write your answer here
```



`read_delim()` execution is reporting the dimensions of the file, along with the guessed delimiter and data type of each columns

If we are happy with the guessed delimiter and the column names / types, we could silent this reporting.

##### Load again the same file, silencing the `read_delim()` message

```{r}
# Write your answer here
```


```{r}
# Write your answer here
```


## The tibble

`read_delim()` loads the data as a tibble. The main advantage to use tibbles over a regular data frame is the printing.

- Tibbles show some useful information such as the number of rows and columns:
    + Look at the top of the tibble and find the information "A tibble rows x cols"
    + How many rows are in the tibble?
- The columns of a tibble report their type:
    + Look at the tibble header, the type of a columns is reported just below its name.
    + What is the type of the `age` column?
    
```{r}
# Write your answer here
```



Actually, both `age` and `id` are **integers**, and should be read as such.

##### Read the `blood_fat.csv` specifying the data types of `age` and `id` as integers


```{block, opts.label = "tip"}
In the `col_types = c(....)` you can use the columns bare names and either the long description
 to call the specific data type like `col_integer()` or the shortcut `"i"`
```

```{r}
# Write your answer here
```


```{r}
# Write your answer here
```


##### Read the `blood_fat.csv` specifying the data types of `age` and `id` as integers, skipping `weight`

```{r}
# Write your answer here
```


```{r}
# Write your answer here
```


##### One summarisation: compute the **mean** of both the `age` and `weight` per `group` {.bonus}

```{r}
# Write your answer here
```

