---
title: "My answers"
author: "My name"
date: "2023-02-01"
output: html_document
---




> This little tutorial aims to make you familiar with some of the functions of the `stringr` package and a few regular expressions. 



## Strings and escape sequences in R

##### Write a sentence with escape sequences.
Try the sentence: `"It's the end of the world!" he said.\` . 
Assign the string to a variable and try as`print()`, `cat()` and `writeLines()`.

```{r}
# Write your answer here
```


## `stringr` functions 

We will be using the `words` data that is built into `stringr`. 
The data set is available to you if you load the package.

```{r}
length(words)
```


##### Select words that contain a `y`

```{r}
# Write your answer here
```


##### Retrieve a boolean vector that indicates which words start with `y`

```{r}
# Write your answer here
```


##### Retrieve the indeces for all words containing `ch` within the word. 

`ch` must not be present at the start or the end of the word. 

```{r}
# Write your answer here
```


##### Extract the `y` and the previous character.

Note: Use the function `unique()` around the results to avoid printing many empty matches.

```{r}
# Write your answer here
```



##### Get the lengths of the first ten words

You can use `head(words, 10)` as a convenient way to access the elements of the `words` vector.
```{r}
# Write your answer here
```



## Viral research 

Read the genome sequence of the Hepatitis D virus: [hepd.fasta](https://biostat2.uni.lu/practicals/data/hepd.fasta).
For now, just execute the following:

```{r}
hepd <- readr::read_lines("https://biostat2.uni.lu/practicals/data/hepd.fasta")
```

##### What is the length of the genome sequence?

```{r}
# Write your answer here
```


##### What is the sequence composition? How often does each character occur?

```{r}
# Write your answer here
```



##### Find the positions of motifs in the sequence  

Find all matches of the sequence $ATG$ in the sequence. 

```{r}
# Write your answer here
```



##### Find motifs in the sequence using `str_locate()`. 



## Tidyr `separate()`

##### Use `tidyr`'s separate function

Create the `dirty_frame` as below. Then split the `data` column into a `name` ,  `age` and `wealth` column.  

There are other options to split the columns. This exercise is optional.


```{r}
dirty_frame <- tibble(id = seq(1, 5),
                      data = c( "as-18-123", "bea-22-122.4", "d:9:201", 'fred-23-222', 'gunter=58=56'))
```


```{r}
#| solution: true 
separate_wider_regex(dirty_frame, data, c(name   = "\\w+", 
                                          ".", 
                                          age    = "\\d+",
                                          ".",
                                          wealth = "[\\d\\.]+"))

```

