---
title: "My answers"
author: "My name"
date: "2021-10-22"
output: html_document
---




>This practical connects most lectures and practicals covered in the course as they would work together for a typical data analysis with data import, transformation, summarizing and plotting.

# Carbon dixoide plant intake

The dataset is from the paper by **Potvin, C., Lechowicz, M. J. and Tardif, S.** (1990). The statistical analysis of ecophysiological response curves obtained from experiments involving repeated measures., _Ecology_, **71**, 1389–1400. You can find it at this [URL](https://biostat2.uni.lu/practicals/data/CO2.tsv)

## Load the data

##### Load the [dataset](https://biostat2.uni.lu/practicals/data/CO2.tsv), assign the name `plant_CO2`

```{block, opts.label = "tip"}
Pay attention to use meaningful names. 
`plant_CO2_url` is a good name for the URL of this dataset.
`plant_CO2_file` is appropriate for tsv file if you download it.
```




```{r}
# Write your answer here
```

## Explore the data

##### How many rows and columns are present? And which data types are they?

```{r}
# Write your answer here
```


##### Three columns are categorical, taken together, how many observations have they?

```{block, opts.label = "tip"}
Using `count()` with the relevant columns is the easiest approach.
Sorting by the highest number is often a nice way to discover the data.
This option is available in `count()`
```

```{r}
# Write your answer here
```


```{r}
# Write your answer here
```


##### How many categories for each of the character columns do you find?

```{block, opts.label = "tip"}
One option is to find the number of unique values using `n_distinct()`.
This would have to be done three times, one for each character column. 
Otherwise, `n_distinct()` can be applied to character columns using
`across()` and relevant `where()` condition.
```

```{block, opts.label = "warning"}
`n_distinct()`, `across()`  are functions that work **inside** `mutate()` or `summarise()`.
```

```{r}
# Write your answer here
```


##### Rename the columns to make them all lower case and continue with the resulting tibble for the next steps.

```{r}
# Write your answer here
```


## CO<sub>2</sub> data

The numeric columns are defined as:

- `conc` is ambient carbon dioxide concentrations ($mL/L$).
- `uptake` is carbon dioxide uptake rates ($umol/m^2$ sec).

##### Compute the mean of `conc` and `uptake` depending on the location (Quebec or Mississippi)

```{block, opts.label = "tip"}
Ignore the treatment, chilled or / non chilled for now
```

```{r}
# Write your answer here
```


##### Does the CO<sub>2</sub> concentrations appear different in both locations?

```{r}
# Write your answer here
```


##### Compute the mean of `conc` and `uptake` depending on the plant location (`type` column), `plant` and treatment. Add the number of observations per group.

```{block, opts.label = "tip"}
the number of observations should be **7** everywhere as you saw before
```

```{r}
# Write your answer here
```


##### From the summary above, sort the table to show the plant / location / treatment that uptake the most carbon dioxide

```{block, opts.label = "tip"}
a descending ordering is possible with `desc(mean_uptake)` in the `arrange()` function
```


```{r}
# Write your answer here
```


It is hard to see if all plants have both treatment, you can pivot the table to find out. Right now the table after summarisation is in the **long** format. Pivoting means converting the table to the **wide format** using the treatment unique values (**chilled** and **nonchilled**) as new columns and filled with the values taken from the `mean_uptake` column.

##### Pivot the table using `plant` and `type` as ids, names from `treatment` and values from the mean of carbon dioxide uptake

```{block, opts.label = "tip"}
look up the different arguments of the function `tidyr::pivot_wider()`
```


```{r}
# Write your answer here
```


##### What do you conclude, does the same plant have been assessed for both treatment?

```{r}
# Write your answer here
```


## Visualize the location and CO<sub>2</sub> uptake

Numbers are nice but plotting the data might better reveal how location and treatment relate.

##### Plot the CO<sub>2</sub> uptake in function of plant location, using violin plots filled by treatment {.bonus}

```{r}
# Write your answer here
```


As a motivation, here is one version of the plot:



